Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2013-03-22 documents/mixtures/response_to_referee.txt

   This paper gives a particularly lucid, concise and clear
   description of a relatively old idea. As such, it represents a
   significant contribution to the literature as a pedagogical
   reference. It should be published but see 4. below.

We agree with the referee that the idea of using Gaussians to fit
galaxies is old but we don't agree with any implication that the
method or applied math results shown are old.  All references we can
find in the literature show *how* to do mixture-of-Gaussian fitting in
real imaging, but don't give mixture-of-Gaussian approximations to
known profiles.  This paper, in that sense, is more of an "applied
math meets astronomy" paper than the previous work.  We will return to
this below, but we *did* make changes to the text in response to this
general concern of the referee: We added sentences in the introductory
paragraphs and at the end of the paper to emphasize the difficulty of
rendering the profiles without these approximations.

To be more specific: we added some sentences to emphasize that the
currently used methods for rendering profiles are effectively mixtures
of delta functions (the profile is sampled at a finite set of
locations) and rendering requires as many PSF renderings as there are
sample points.  This is a huge computation.

We also added an introductory paragraph that points out that the code
complexity required to do straight profile rendering (ie without the
MoGs) correctly is large and this leads to engineering overheads in
testing and correctness that go way beyond the simple numerical speed
losses.  We also bring up the point that conventional numerical
integration methods effectively treat the profiles as very complex
mixtures of kernels.

@@

   Equation 19: I did not quite understand your distinction between
   the \Delta X vector here and the x vector in equation 2. Please
   clarify.

Agreed.  We added three sentences shortly after equation (19)
explaining explicitly why we use \Delta x and not x here.  It relates
to the fact that the PSF is only used as a convolution kernel, never
as a model in itself.

In addition to the changes given above, we also made the following
small changes:

* Added references to Kochanek 2000 and van den Bosch 2008 and
  Bendinelli 1993 and Spergel 2010.

* Fixed a few typos.

* Added some acknowledgements, including one to the referee.

### 2013-03-22 documents/mixtures/mixture_models.tex
em-dash between K&L in K--L div; alsoo -> also

We seek the \emph{best} MoG approximations.  This necessitates
definition of the word ``best''.  If we think of the profiles as being
two-dimensional probability distribution functions (for, say, the
arrivals of photons), then one natural choice is the Kullback-Leibler
divergence or similar cross-entropy or information-theoretic measure.
However, in typical astronomical imaging, the galaxy is superimposed
on a substantial, flat sky level, and the noise in the data is close
to Gaussian.  This suggests more chi-squared-like objectives.  We
adopt the latter, in part because they are most appropriate for our
specific proposed application (modeling \sdss-like astronomical
imaging), but experiments we have performed suggest that
information-theoretic objectives also lead to good results.  In
principle the objective that should be used to make these
approximations could be slightly different in different contexts.  In
practice, since the dependence of ``best objective'' on data
properties is weak and because the point here is to make generic
replacements for the analytic profiles, chi-squared is a good choice.

>>>>

We seek the \emph{best} MoG approximations.  This necessitates
definition of the word ``best''.  If we think of the profiles as being
two-dimensional probability distribution functions (for, say, the
arrivals of photons), then one natural choice is the Kullback--Leibler
divergence or similar cross-entropy or information-theoretic measure.
However, in typical astronomical imaging, the galaxy is superimposed
on a substantial, flat sky level, and the noise in the data is close
to Gaussian.  This suggests more chi-squared-like objectives.  We
adopt the latter, in part because they are most appropriate for our
specific proposed application (modeling \sdss-like astronomical
imaging), but experiments we have performed suggest that
information-theoretic objectives also lead to good results.  In
principle the objective that should be used to make these
approximations could be slightly different in different contexts.  In
practice, since the dependence of ``best objective'' on data
properties is weak and because the point here is to make generic
replacements for the analytic profiles, chi-squared is a good choice.
